---
author: "Anıl İsmet AŞCI"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
# libraries ---------------------------------------------------------------------------------------------------

if (!require(rvest)) install.packages('rvest')
if (!require(ggplot2)) install.packages('ggplot2')
if (!require(mice)) install.packages('mice')
if (!require(countrycode)) install.packages('countrycode')
if (!require(data.table)) install.packages('data.table')
if (!require(rgdal)) install.packages('rgdal')
if (!require(plotly)) install.packages('plotly')
if (!require(rgdal)) install.packages('rgdal')
if (!require(dplyr)) install.packages('dplyr')
if (!require(plotly)) install.packages('plotly')


library(rgdal)
library(rvest)
library(ggplot2)
library(mice)
library(countrycode)
library(data.table)
library(plotly)
library(dplyr)
library(rgdal)
library(plotly)

# data --------------------------------------------------------------------------------------------------------

#Life expectancy source
url<- c("https://en.wikipedia.org/wiki/List_of_countries_by_life_expectancy")
fp <- read_html(url)
str(fp)


##required html class for table
tables<- html_nodes(fp, ".mw-datatable")
class(tables)

## assigning tables to individual table, table 1 creates problem so do not assign it

for (x in c(2:length(tables))) {
  assign(paste0("table",x), html_table(tables[x], header = T, dec = ".", fill=TRUE), envir = .GlobalEnv)
}

## To names data according to institutions rerun the tables and look at the datatable names
tables


## Tables and their content
WHO2015<-table2[[1]]
UN2015 <- table3[[1]]
GBD2010 <- table4[[1]]
CIA2017 <- table5[[1]]
OECD2016 <- table6[[1]]


## I will choose CIA2017 since it has the max. elements and it is rather new
str(CIA2017)
class(CIA2017)
setDT(CIA2017)
setorder(CIA2017,"Country")
## Check for NA
md.pattern(CIA2017)
CIA2017 <- na.omit(CIA2017)
md.pattern(CIA2017)



CIA2017[, difference := Female-Male]

## Graph the CIA2017
names(CIA2017)


ggplot(CIA2017, aes(x=Male,y=Female,colour= Country))+geom_point()+theme(legend.position = "none")+
  stat_smooth(method = 'lm', aes(colour = 'linear'), se = FALSE)




# Education index source

url2<- c("https://en.wikipedia.org/wiki/Education_Index")
fp2 <- read_html(url2)
str(fp2)
fp2

## required html class for table
tables2<- html_nodes(fp2, ".wikitable")
class(tables2)
tables2



## Assigning table to required table

for (x in c(1:length(tables2))) {
  assign(paste0("table2",x), html_table(tables2[x], header = T, dec = ".", fill=TRUE), envir = .GlobalEnv)
}

## Since there is only one table that we could use. I choosed that
data21 <- table21[[1]]
##  names in data21
names(data21)
## re-name it as edu and do the neccesary work
edu <- data21
nrow(edu)
class(edu)
edu<- data.table(edu)
setorder(edu,Country)

## Check for NA
md.pattern(edu)


## Graph the CIA2017
names(edu)

### rename the columns for ease

colnames(edu) <- c("Rank",
                   "Country",
                   "EduIndex",
                   "ExpSch",
                   "MSch",
                   "HDI",
                   "Continent")


ggplot(edu, aes(x=ExpSch,y=MSch,colour= Country))+geom_point()+theme(legend.position = "none")+
  stat_smooth(method = 'lm', aes(colour = 'linear'), se = FALSE)


## Combining two datasets


nrow(CIA2017)-nrow(edu)
#### There are extra countries in CIA2017 datatable. 
#### Let's check if country names are same for both  datatables

educ <- edu$Country
ciac <- CIA2017$Country


a <- c()
b <- c()


### Countries in edu but not in CIA
for (x in educ){
  if (sum(x == CIA2017$Country) == 0) {
    a[x] <- x
  }
  
}
a

### Countries in CIA but not in edu
  
for (x in ciac){
  if (sum(x == edu$Country) == 0) {
   b[x] <- x
  
  }
  
}

b


length(a)
length(b)

length(b)-length(a)

#### There are 19 countries both exists in two datasets but named differently,
#### We had also found the 36 countries not exists in edu. 
#### By using countrycode package lets add iso codes of countries then merge it

edu[, iso2c := countrycode(Country, origin ='country.name', destination ='iso2c')]
CIA2017[, iso2c := countrycode(Country, origin ='country.name', destination ='iso2c')]
edu<-na.omit(edu)
CIA2017<- na.omit(CIA2017)
datum<-merge(CIA2017,edu , by="iso2c")
nrow(datum)
nrow(CIA2017)-36

#### It is similar
### Lets check if there is any NA
md.pattern(datum)
class(datum)
names(datum)

datum

datum[, Country.y := NULL]
names(datum)
colnames(datum)<- c("iso2c" ,
                    "Country",
                    "Male",
                    "Female",
                    "Both",
                    "difference",
                    "difference2",
                    "Rank",
                    "EduIndex",
                    "ExpSch",
                    "MSch",
                    "HDI",
                    "Continent")
str(datum)

### Let's graph it for test
ggplot(datum, aes(x=MSch,y=Both,colour= Country))+geom_point()+theme(legend.position = "none")+
  stat_smooth(method = 'lm', aes(colour = 'linear'), se = FALSE)

datum[,mean(Both), by= Continent]

# theme for graphs ---------------------------------------------------------------------------
theme_ismetovic<-function (base_size = 12, base_family = "", quon = 0.9){
  half_line <- base_size/2
  theme(
    rect = element_blank(),
    text = element_text(family = base_family, face = "plain",
                        colour = "#4D4845", size = base_size,
                        lineheight = 1,  hjust = 0.5,
                        vjust = 0.5, angle = 0, 
                        margin = margin(), debug = FALSE),
    
    #plot  
    plot.background = element_blank(), 
    plot.title = element_text(face = "bold",size = rel(1.2),hjust = 0, margin = margin(b = half_line )),
    plot.subtitle = element_text(size= rel(0.9),hjust = 0, margin= margin(b = half_line )),
    plot.caption = element_text(face = "bold",color = "#4D4845", hjust = 0,  size = rel(0.8), margin = margin(t = half_line) ),
    #axis  
    axis.text = element_text(size = rel(quon), colour = "#4D4845"),
    axis.text.x = element_text(margin = margin(t = 0.8*half_line/2), vjust = 1), 
    axis.text.y = element_text(margin = margin(r = 0.8*half_line/2), hjust = 1),
    axis.title = element_text(color =  "#4D4845",size = rel(1),face = "bold"),
    
    #legend
    
    legend.key.size = unit(2, "line"), 
    legend.text = element_text(size = rel(0.9), color = "#4D4845" ),
    legend.title = element_text(hjust = 0, color ="#4D4845" , size = rel(0.9),face = "bold"), 
    legend.position = "top",
    legend.title.align = 0  , 
    legend.direction = "horizontal",
    legend.justification = c(0,0), 
    
    #panel  
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color="#F2E5DA", size = 0.2),
    panel.margin.y = NULL, panel.ontop = FALSE, 
    
    #strip in case of facet_wrap    
    strip.background = element_rect(fill="white"),
    strip.text = element_text(colour = "#4D4845", size = rel(0.8)),
    
    complete = TRUE)
}


# Graphs for data -------------------------------------------------------------------------------------


a1 <- ggplot(datum,aes(x=MSch,y=Both))+
  geom_point()+
  geom_smooth(method=lm , color="red", se=FALSE)+
  theme_ismetovic()+
  labs(title = "Schooling and Expected Life from Birth",
       subtitle = "",
       x = "Mean years of schooling", y = "Expected life from birth")
a1

a2 <- ggplot(datum,aes(x=MSch,y=Both,color= Continent,group=Country))+
  geom_point()+
  theme_ismetovic()+
  labs(title = "Schooling and Expected Life from Birth",
       subtitle = "",
       x = "Mean years of schooling", y = "Expected life from birth")
a2
ggplotly(a2)

expp<-expression('45'~degree)

a3 <- ggplot(datum,aes(x=Male,y=Female))+
  geom_point()+
  geom_abline()+
  theme_ismetovic()+
  labs(title = "Expected Life from Birth",
       subtitle = "",
       x = "Males", y = "Females")+
  annotate(geom="text", x=85, y= 90, label=expp,color="black")
a3

a4 <- ggplot(datum,aes(x=MSch,y=difference,color= Continent,group=Country))+
  geom_point()+
  theme_ismetovic()+
  labs(title = "Expected Life from Birth",
       subtitle = "Difference between genders",
       x = "Mean years of schooling", y = "Difference")+
  geom_hline(yintercept = 0)
a4

ggplotly(a4)


# World Graphs for data---------------------------------------------------------------------------
#### Geospatial data used is from Erol Taymaz

# Downloading the Map File, data will be avalible as long as the metu 

wa <- readOGR("C:/Users/aiasc/OneDrive - metu.edu.tr/desktop/R","ne_10m_admin_0_countries")
 
# Deleting antartica and creating dataset for further use and testing on graph

wa <- subset(wa, admin!="Antarctica")
wf <- fortify(wa)
wa$id <- row.names(wa)
wf <- dplyr::left_join(wf, wa@data)
names(wf)
wf <- data.table(wf)
wf
a<-ggplot(wf, aes(long, lat, group = group, fill=gdp_md_est )) + 
  geom_polygon()

a

### adding life  expectency and education data into wf
### Each Country have multiple elements due to their shapes
### Result of that, for one element in datum we need to create as many element as needed


############################################################################################

#----------------Attention!. Each one of the variables may require 10 min.----------------#

#----------Please Notify me if you know a shorter way via "aiasci@outlook.com" -----------#

############################################################################################

qw<-wf$iso_a2

# loop for getting Both life, test whether the length of the created list equals number of rows, then add it
life<-c()
for (ii in 1:length(qw)) {
  if (sum(qw[ii] == datum$iso2c) == 1) {
    result <- datum[iso2c == qw[ii]]$Both
    life[ii] <- result
    
  }
  else {
    life[ii] <- 0
  }
  
}


length(life)
nrow(wf)- length(life)
wf[, life := life]

# loop for getting Female life, test whether the length of the created list equals number of rows, then add it
flife<-c()
for (ii in 1:length(qw)) {
  if (sum(qw[ii] == datum$iso2c) == 1) {
    result <- datum[iso2c == qw[ii]]$Female
    flife[ii] <- result
    
  }
  else {
    flife[ii] <- 0
  }
  
}

length(flife)
nrow(wf)- length(flife)
wf[, flife := flife]

# loop for getting Male life, test whether the length of the created list equals number of rows, then add it
mlife<-c()
for (ii in 1:length(qw)) {
  if (sum(qw[ii] == datum$iso2c) == 1) {
    result <- datum[iso2c == qw[ii]]$Male
    mlife[ii] <- result
    
  }
  else {
    mlife[ii] <- 0
  }
  
}

length(mlife)
nrow(wf)- length(mlife)
wf[, mlife := mlife]

# loop for getting Expected year of schooling life,  test whether the length of the created list equals number of rows, then add it
Esch<-c()
for (ii in 1:length(qw)) {
  if (sum(qw[ii] == datum$iso2c) == 1) {
    result <- datum[iso2c == qw[ii]]$ExpSch
    Esch[ii] <- result
    
  }
  else {
    Esch[ii] <- 0
  }
  
}

length(Esch)
nrow(wf)- length(Esch)
wf[, Esch := Esch]


# loop for getting Mean year of schooling life,  test whether the length of the created list equals number of rows, then add it
Msch<-c()
for (ii in 1:length(qw)) {
  if (sum(qw[ii] == datum$iso2c) == 1) {
    result <- datum[iso2c == qw[ii]]$MSch
    Msch[ii] <- result
    
  }
  else {
    Msch[ii] <- 0
  }
  
}

length(Msch)
nrow(wf)- length(Msch)
wf[, Msch := Msch]

names(wf)

## Graphs

wf[, lifeg := ifelse(life>=80, "80 \u2265 ",
                     ifelse(life>=70 & life <80 , "80> \u2265 70",
                            ifelse(life==0,"No Data" ,       
                                   ifelse(life>60 & life <70, "70> \u2265 60","60<" ))))]

a11<-ggplot(wf, aes(long, lat, group = group, fill = lifeg)) + 
  geom_polygon() + 
  coord_equal(ratio = 1.3, xlim = NULL, ylim = NULL, expand = T)+
  theme_ismetovic()+
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
       axis.text.y=element_blank(),axis.ticks.y=element_blank(),legend.title = element_blank())+
  scale_fill_manual(values = c("#bdc9e1","#74a9cf","#2b8cbe","#045a8d","#f1eef6"))+
  labs(title = "Expected life from Birth",
       subtitle = "CIA Factbook 2017")

wf[, mschg := ifelse(Msch>=11, "11 \u2265 ",
                     ifelse(Msch>=8 & Msch <11 , "11> \u2265 8",
                            ifelse(Msch==0,"No Data" ,       
                                   ifelse(Msch>80 & Msch <5, "8> \u2265 5","5<" ))))]

a12<-ggplot(wf, aes(long, lat, group = group, fill = mschg)) + 
  geom_polygon() + 
  coord_equal(ratio = 1.3, xlim = NULL, ylim = NULL, expand = T)+
  theme_ismetovic()+
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),axis.ticks.y=element_blank(),legend.title = element_blank())+
  scale_fill_manual(values = c("#d7301f","#fc8d59","#fdcc8a","#fef0d9"))+
  labs(title = "Mean years of Education",
       subtitle = "for People ages 25 and over")


wf[, eschg := ifelse(Esch>=16, "16 \u2265 ",
                     ifelse(Esch>=12 & Esch <16 , "16> \u2265 12",
                            ifelse(Msch==0,"No Data" ,       
                                   ifelse(Esch>=8& Esch <12, "12> \u2265 8 ","8<" ))))]

a13<-ggplot(wf, aes(long, lat, group = group, fill = eschg)) + 
  geom_polygon() + 
  coord_equal(ratio = 1.3, xlim = NULL, ylim = NULL, expand = T)+
  theme_ismetovic()+
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),axis.ticks.y=element_blank(),legend.title = element_blank())+
  scale_fill_manual(values = c("#f768a1","#7a0177","#c51b8a","#fbb4b9","#feebe2"))+
  labs(title = "Expected years of Education",
       subtitle = "for People under 18")


# Lets Arrange them for tables ------------------------------------------------------------------------------------

datum_m <- datum[,c(1:3,12)]
datum_f <- datum[,c(1:2,4,12)]
datum_b <- datum[,c(1:2,5,12)]
datum_diff <- datum[,c(1:2,6,12)]
datum_eduindex <- datum[,c(1:2,8,12)]
datum_expsch <- datum[,c(1:2,9,12)]
datum_msch <- datum[,c(1:2,10,12)]


datum_m[, tip := rep("Expected life from birth for Males", nrow(datum_m))]
datum_f[, tip := rep("Expected life from birth for Females", nrow(datum_f))]
datum_diff[, tip := rep("Difference between Females and Males", nrow(datum_diff))]
datum_b[, tip := rep("Expected life from birth", nrow(datum_b))]
datum_msch[, tip := rep("Mean years of schooling", nrow(datum_msch))]
datum_expsch[, tip := rep("Expected years of schooling", nrow(datum_expsch))]
datum_eduindex[, tip := rep("Education Index", nrow(datum_eduindex))]

colnamesaremess <- c("iso2c","Country","Result","Continent","tip")
colnames(datum_m) <- colnamesaremess
colnames(datum_f) <- colnamesaremess
colnames(datum_b) <- colnamesaremess
colnames(datum_diff) <- colnamesaremess
colnames(datum_msch) <- colnamesaremess
colnames(datum_expsch) <- colnamesaremess
colnames(datum_eduindex) <- colnamesaremess

setorder(datum_m, -Result)
setorder(datum_f, -Result)
setorder(datum_b, -Result)
setorder(datum_diff, -Result)
datum_diff$Result <- round(datum_diff$Result,2)
setorder(datum_msch, -Result)
setorder(datum_expsch, -Result)
setorder(datum_eduindex, -Result)

datum_life <- rbind(datum_m,datum_f,datum_b)


datum_edu <- rbind(datum_expsch,datum_msch)

datum_diff 

datum_eduindex

# For general statistics part these codes will create the tables.

DT::datatable(datum_b)
DT::datatable(datum_f)
DT::datatable(datum_m)
DT::datatable(datum_diff)
DT::datatable(datum_msch)
DT::datatable(datum_expsch)
DT::datatable(datum_eduindex)




```


# Life expectancy and education
<div style="text-align: justify;font-size: 15px;">  
It is a known fact that education and life expectancy are positively correlated since with the education, productivity and corollary income of the population increases. Also, higher years of education decreases criminality. Thus the increase in the years' of compulsory education eventually has a positive effect on the life expectancy.
</div> 
<br>
<div style="text-align: justify;font-size: 15px;">  
For this homework, I scrap data from Wikipedia's [Life Expectancy](https://en.wikipedia.org/wiki/List_of_countries_by_life_expectancy) and [Education Index](https://en.wikipedia.org/wiki/Education_Index) via required R packages. I used the CIA's 2017 Life Expectancy data and 2016 Human Development Report's Education Index embedded in the website. These data sets do not have a time series; thus, I will only make comparisons. For further examination, tables below can be used. 
</div> 
<br>
<div style="text-align: justify;font-size: 15px;">  
Lastly data for maps is taken from Erol Taymaz.
</div> 

## Graphs {.tabset}
### Schooling on Life Expectancy {.tabset .tabset-fade}
<div style="text-align: justify;font-size: 15px;">  
It is obvious that mean year of schooling for people ages 25 and older is positively correlated with life expectancy.
</div> 
```{r echo=FALSE, warning=FALSE ,message=FALSE, out.width= '120%'}
a1
```

### Schooling on Life Expectancy (interactive) {.tabset .tabset-fade}
<div style="text-align: justify;font-size: 15px;">  
Again as expected countries are underdeveloped, in war and experienced recent disasters are at the bottom-left of the table while the so-called prosperous ones are at the top-right. 
</div> 
```{r echo=FALSE, warning=FALSE ,message=FALSE, out.width= '120%'}
ggplotly(a2)
```

### Male-Female difference {.tabset .tabset-fade}
<div style="text-align: justify;font-size: 15px;">  
Also, females expected to live longer. 
</div> 
```{r echo=FALSE, warning=FALSE ,message=FALSE, out.width= '120%'}
a3
```

### Difference (interactive) {.tabset .tabset-fade}
<div style="text-align: justify;font-size: 15px;">  
The expected years of living difference between genders shows that old USSR countries absurdly on the top while in Eswatini males expected to live 1.2 years more than females. 
</div> 

```{r echo=FALSE, warning=FALSE , message=FALSE, out.width= '120%'}
ggplotly(a4)
```



## Global Graphs {.tabset}
I do not include plotly based graphs due to high system requirement needs which my computer do not have. Instead, non-interactive plots will be shown.

### Life Expectancy {.tabset .tabset-fade}

```{r echo=F,  out.width= '120%'}

ggplotly(a11)

```

### Mean Years of Education {.tabset .tabset-fade}

```{r echo=F, out.width= '120%'}
ggplotly(a12)
```


### Expected Years of Education {.tabset .tabset-fade}

```{r echo=F, out.width= '120%'}
ggplotly(a13)
```


## General Statistics {.tabset}



### Expected Lifes from Birth {.tabset .tabset-fade}
#### All {.tabset .tabset-fade}
```{r echo=F}

DT::datatable(datum_b)

```

#### Females {.tabset .tabset-fade}
```{r echo=F}

DT::datatable(datum_f)

```

#### Males {.tabset .tabset-fade}
```{r echo=F}

DT::datatable(datum_m)

```

#### Difference {.tabset .tabset-fade}
```{r echo=F}

DT::datatable(datum_diff)

```

### Education {.tabset .tabset-fade}

#### Mean Years {.tabset .tabset-fade}
```{r echo=F}

DT::datatable(datum_msch)

```

#### Expected Years {.tabset .tabset-fade}
```{r echo=F}

DT::datatable(datum_expsch)

```

#### Education Index{.tabset .tabset-fade}
```{r echo=F}

DT::datatable(datum_eduindex)

```



## Appendix

```{r eval=FALSE, warning=FALSE, message= FALSE}
# libraries ---------------------------------------------------------------------------------------------------

if (!require(rvest)) install.packages('rvest')
if (!require(ggplot2)) install.packages('ggplot2')
if (!require(mice)) install.packages('mice')
if (!require(countrycode)) install.packages('countrycode')
if (!require(data.table)) install.packages('data.table')
if (!require(rgdal)) install.packages('rgdal')
if (!require(plotly)) install.packages('plotly')
if (!require(rgdal)) install.packages('rgdal')
if (!require(dplyr)) install.packages('dplyr')



library(rgdal)
library(rvest)
library(ggplot2)
library(mice)
library(countrycode)
library(data.table)
library(plotly)
library(dplyr)
library(rgdal)

# data --------------------------------------------------------------------------------------------------------

#Life expectancy source
url<- c("https://en.wikipedia.org/wiki/List_of_countries_by_life_expectancy")
fp <- read_html(url)
str(fp)


##required html class for table
tables<- html_nodes(fp, ".mw-datatable")
class(tables)

## assigning tables to individual table, table 1 creates problem so do not assign it

for (x in c(2:length(tables))) {
  assign(paste0("table",x), html_table(tables[x], header = T, dec = ".", fill=TRUE), envir = .GlobalEnv)
}

## To names data according to institutions rerun the tables and look at the datatable names
tables


## Tables and their content
WHO2015<-table2[[1]]
UN2015 <- table3[[1]]
GBD2010 <- table4[[1]]
CIA2017 <- table5[[1]]
OECD2016 <- table6[[1]]

tablelist <- list(WHO2015,UN2015,GBD2010,CIA2017,OECD2016)

## names in datatables
lapply(tablelist, names)


## setting order for datatables
for (x in tablelist) {
  setorder(x,Country)
}

## nrows for datatables
lapply(tablelist, nrow)

## I will choose CIA2017 since it has the max. elements and it is rather new
str(CIA2017)
class(CIA2017)
CIA2017<- data.table(CIA2017)

## Check for NA
md.pattern(CIA2017)
CIA2017 <- na.omit(CIA2017)
md.pattern(CIA2017)



CIA2017[, difference := Female-Male]

## Graph the CIA2017
names(CIA2017)


ggplot(CIA2017, aes(x=Male,y=Female,colour= Country))+geom_point()+theme(legend.position = "none")+
  stat_smooth(method = 'lm', aes(colour = 'linear'), se = FALSE)




# Education index source

url2<- c("https://en.wikipedia.org/wiki/Education_Index")
fp2 <- read_html(url2)
str(fp2)
fp2

## required html class for table
tables2<- html_nodes(fp2, ".wikitable")
class(tables2)
tables2



## Assigning table to required table

for (x in c(1:length(tables2))) {
  assign(paste0("table2",x), html_table(tables2[x], header = T, dec = ".", fill=TRUE), envir = .GlobalEnv)
}

## Since there is only one table that we could use. I choosed that
data21 <- table21[[1]]
##  names in data21
names(data21)
## re-name it as edu and do the neccesary work
edu <- data21
nrow(edu)
class(edu)
edu<- data.table(edu)
setorder(edu,Country)

## Check for NA
md.pattern(edu)


## Graph the CIA2017
names(edu)

### rename the columns for ease

colnames(edu) <- c("Rank",
                   "Country",
                   "EduIndex",
                   "ExpSch",
                   "MSch",
                   "HDI",
                   "Continent")


ggplot(edu, aes(x=ExpSch,y=MSch,colour= Country))+geom_point()+theme(legend.position = "none")+
  stat_smooth(method = 'lm', aes(colour = 'linear'), se = FALSE)


## Combining two datasets


nrow(CIA2017)-nrow(edu)
#### There are extra countries in CIA2017 datatable. 
#### Let's check if country names are same for both  datatables

educ <- edu$Country
ciac <- CIA2017$Country


a <- c()
b <- c()


### Countries in edu but not in CIA
for (x in educ){
  if (sum(x == CIA2017$Country) == 0) {
    a[x] <- x
  }
  
}
a

### Countries in CIA but not in edu
  
for (x in ciac){
  if (sum(x == edu$Country) == 0) {
   b[x] <- x
  
  }
  
}

b


length(a)
length(b)

length(b)-length(a)

#### There are 19 countries both exists in two datasets but named differently,
#### We had also found the 36 countries not exists in edu. 
#### By using countrycode package lets add iso codes of countries then merge it

edu[, iso2c := countrycode(Country, origin ='country.name', destination ='iso2c')]
CIA2017[, iso2c := countrycode(Country, origin ='country.name', destination ='iso2c')]
edu<-na.omit(edu)
CIA2017<- na.omit(CIA2017)
datum<-merge(CIA2017,edu , by="iso2c")
nrow(datum)
nrow(CIA2017)-36

#### It is similar
### Lets check if there is any NA
md.pattern(datum)
class(datum)
names(datum)

datum

datum[, Country.y := NULL]
names(datum)
colnames(datum)<- c("iso2c" ,
                    "Country",
                    "Male",
                    "Female",
                    "Both",
                    "difference",
                    "Rank",
                    "EduIndex",
                    "ExpSch",
                    "MSch",
                    "HDI",
                    "Continent")
str(datum)

### Let's graph it for test
ggplot(datum, aes(x=MSch,y=Both,colour= Country))+geom_point()+theme(legend.position = "none")+
  stat_smooth(method = 'lm', aes(colour = 'linear'), se = FALSE)

datum[,mean(Both), by= Continent]

# theme for graphs ---------------------------------------------------------------------------
theme_ismetovic<-function (base_size = 12, base_family = "", quon = 0.9){
  half_line <- base_size/2
  theme(
    rect = element_blank(),
    text = element_text(family = base_family, face = "plain",
                        colour = "#4D4845", size = base_size,
                        lineheight = 1,  hjust = 0.5,
                        vjust = 0.5, angle = 0, 
                        margin = margin(), debug = FALSE),
    
    #plot  
    plot.background = element_blank(), 
    plot.title = element_text(face = "bold",size = rel(1.2),hjust = 0, margin = margin(b = half_line )),
    plot.subtitle = element_text(size= rel(0.9),hjust = 0, margin= margin(b = half_line )),
    plot.caption = element_text(face = "bold",color = "#4D4845", hjust = 0,  size = rel(0.8), margin = margin(t = half_line) ),
    #axis  
    axis.text = element_text(size = rel(quon), colour = "#4D4845"),
    axis.text.x = element_text(margin = margin(t = 0.8*half_line/2), vjust = 1), 
    axis.text.y = element_text(margin = margin(r = 0.8*half_line/2), hjust = 1),
    axis.title = element_text(color =  "#4D4845",size = rel(1),face = "bold"),
    
    #legend
    
    legend.key.size = unit(2, "line"), 
    legend.text = element_text(size = rel(0.9), color = "#4D4845" ),
    legend.title = element_text(hjust = 0, color ="#4D4845" , size = rel(0.9),face = "bold"), 
    legend.position = "top",
    legend.title.align = 0  , 
    legend.direction = "horizontal",
    legend.justification = c(0,0), 
    
    #panel  
    panel.background = element_blank(),
    panel.grid.major.y = element_line(color="#F2E5DA", size = 0.2),
    panel.margin.y = NULL, panel.ontop = FALSE, 
    
    #strip in case of facet_wrap    
    strip.background = element_rect(fill="white"),
    strip.text = element_text(colour = "#4D4845", size = rel(0.8)),
    
    complete = TRUE)
}


# Graphs for data -------------------------------------------------------------------------------------


a1 <- ggplot(datum,aes(x=MSch,y=Both))+
  geom_point()+
  geom_smooth(method=lm , color="red", se=FALSE)+
  theme_ismetovic()+
  labs(title = "Schooling and Expected Life from Birth",
       subtitle = "",
       x = "Mean years of schooling", y = "Expected life from birth")
a1

a2 <- ggplot(datum,aes(x=MSch,y=Both,color= Continent,group=Country))+
  geom_point()+
  theme_ismetovic()+
  labs(title = "Schooling and Expected Life from Birth",
       subtitle = "",
       x = "Mean years of schooling", y = "Expected life from birth")
a2
ggplotly(a2)

expp<-expression('45'~degree)

a3 <- ggplot(datum,aes(x=Male,y=Female))+
  geom_point()+
  geom_abline()+
  theme_ismetovic()+
  labs(title = "Expected Life from Birth",
       subtitle = "",
       x = "Males", y = "Females")+
  annotate(geom="text", x=85, y= 90, label=expp,color="black")
a3

a4 <- ggplot(datum,aes(x=MSch,y=difference,color= Continent,group=Country))+
  geom_point()+
  theme_ismetovic()+
  labs(title = "Expected Life from Birth",
       subtitle = "Difference between genders",
       x = "Mean years of schooling", y = "Difference")+
  geom_hline(yintercept = 0)
a4

ggplotly(a4)


# World Graphs for data---------------------------------------------------------------------------
#### Geospatial data used is from Erol Taymaz

# Downloading the Map File, data will be avalible as long as the metu 
download.file("http://users.metu.edu.tr/e214755/map.zip" , destfile=paste0(getwd(), "map.zip") )
system( paste0("unzip", getwd(), "/map.zip"))
wa <- readOGR(dsn= paste0(getwd(), "/ne_10m_admin_0_countries.shp"))

# Deleting antartica and creating dataset for further use and testing on graph

wa <- subset(wa, admin!="Antarctica")
wf <- fortify(wa)
wa$id <- row.names(wa)
wf <- dplyr::left_join(wf, wa@data)
names(wf)
wf <- data.table(wf)
wf
a<-ggplot(wf, aes(long, lat, group = group, fill=gdp_md_est )) + 
  geom_polygon()

a

### adding life  expectency and education data into wf
### Each Country have multiple elements due to their shapes
### Result of that, for one element in datum we need to create as many element as needed


############################################################################################

#----------------Attention!. Each one of the variables may require 10 min.----------------#

#----------Please Notify me if you know a shorter way via "aiasci@outlook.com" -----------#

############################################################################################

qw<-wf$iso_a2

# loop for getting Both life, test whether the length of the created list equals number of rows, then add it
life<-c()
for (ii in 1:length(qw)) {
  if (sum(qw[ii] == datum$iso2c) == 1) {
    result <- datum[iso2c == qw[ii]]$Both
    life[ii] <- result
    
  }
  else {
    life[ii] <- 0
  }
  
}


length(life)
nrow(wf)- length(life)
wf[, life := life]

# loop for getting Female life, test whether the length of the created list equals number of rows, then add it
flife<-c()
for (ii in 1:length(qw)) {
  if (sum(qw[ii] == datum$iso2c) == 1) {
    result <- datum[iso2c == qw[ii]]$Female
    flife[ii] <- result
    
  }
  else {
    flife[ii] <- 0
  }
  
}

length(flife)
nrow(wf)- length(flife)
wf[, flife := flife]

# loop for getting Male life, test whether the length of the created list equals number of rows, then add it
mlife<-c()
for (ii in 1:length(qw)) {
  if (sum(qw[ii] == datum$iso2c) == 1) {
    result <- datum[iso2c == qw[ii]]$Male
    mlife[ii] <- result
    
  }
  else {
    mlife[ii] <- 0
  }
  
}

length(mlife)
nrow(wf)- length(mlife)
wf[, mlife := mlife]

# loop for getting Expected year of schooling life,  test whether the length of the created list equals number of rows, then add it
Esch<-c()
for (ii in 1:length(qw)) {
  if (sum(qw[ii] == datum$iso2c) == 1) {
    result <- datum[iso2c == qw[ii]]$ExpSch
    Esch[ii] <- result
    
  }
  else {
    Esch[ii] <- 0
  }
  
}

length(Esch)
nrow(wf)- length(Esch)
wf[, Esch := Esch]


# loop for getting Mean year of schooling life,  test whether the length of the created list equals number of rows, then add it
Msch<-c()
for (ii in 1:length(qw)) {
  if (sum(qw[ii] == datum$iso2c) == 1) {
    result <- datum[iso2c == qw[ii]]$MSch
    Msch[ii] <- result
    
  }
  else {
    Msch[ii] <- 0
  }
  
}

length(Msch)
nrow(wf)- length(Msch)
wf[, Msch := Msch]

names(wf)

## Graphs

wf[, lifeg := ifelse(life>=80, "80 \u2265 ",
                     ifelse(life>=70 & life <80 , "80> \u2265 70",
                            ifelse(life==0,"No Data" ,       
                                   ifelse(life>60 & life <70, "70> \u2265 60","60<" ))))]

a11<-ggplot(wf, aes(long, lat, group = group, fill = lifeg)) + 
  geom_polygon() + 
  coord_equal(ratio = 1.3, xlim = NULL, ylim = NULL, expand = T)+
  theme_ismetovic()+
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
       axis.text.y=element_blank(),axis.ticks.y=element_blank(),legend.title = element_blank())+
  scale_fill_manual(values = c("#bdc9e1","#74a9cf","#2b8cbe","#045a8d","#f1eef6"))+
  labs(title = "Expected life from Birth",
       subtitle = "CIA Factbook 2017")

wf[, mschg := ifelse(Msch>=11, "11 \u2265 ",
                     ifelse(Msch>=8 & Msch <11 , "11> \u2265 8",
                            ifelse(Msch==0,"No Data" ,       
                                   ifelse(Msch>80 & Msch <5, "8> \u2265 5","5<" ))))]

a12<-ggplot(wf, aes(long, lat, group = group, fill = mschg)) + 
  geom_polygon() + 
  coord_equal(ratio = 1.3, xlim = NULL, ylim = NULL, expand = T)+
  theme_ismetovic()+
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),axis.ticks.y=element_blank(),legend.title = element_blank())+
  scale_fill_manual(values = c("#d7301f","#fc8d59","#fdcc8a","#fef0d9"))+
  labs(title = "Mean years of Education",
       subtitle = "for People ages 25 and over")


wf[, eschg := ifelse(Esch>=16, "16 \u2265 ",
                     ifelse(Esch>=12 & Esch <16 , "16> \u2265 12",
                            ifelse(Msch==0,"No Data" ,       
                                   ifelse(Esch>=8& Esch <12, "12> \u2265 8 ","8<" ))))]

a13<-ggplot(wf, aes(long, lat, group = group, fill = eschg)) + 
  geom_polygon() + 
  coord_equal(ratio = 1.3, xlim = NULL, ylim = NULL, expand = T)+
  theme_ismetovic()+
  theme(axis.title.x=element_blank(),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),axis.title.y=element_blank(),
        axis.text.y=element_blank(),axis.ticks.y=element_blank(),legend.title = element_blank())+
  scale_fill_manual(values = c("#f768a1","#7a0177","#c51b8a","#fbb4b9","#feebe2"))+
  labs(title = "Expected years of Education",
       subtitle = "for People under 18")


# Lets Arrange them for tables ------------------------------------------------------------------------------------

datum_m <- datum[,c(1:3,12)]
datum_f <- datum[,c(1:2,4,12)]
datum_b <- datum[,c(1:2,5,12)]
datum_diff <- datum[,c(1:2,6,12)]
datum_eduindex <- datum[,c(1:2,8,12)]
datum_expsch <- datum[,c(1:2,9,12)]
datum_msch <- datum[,c(1:2,10,12)]


datum_m[, tip := rep("Expected life from birth for Males", nrow(datum_m))]
datum_f[, tip := rep("Expected life from birth for Females", nrow(datum_f))]
datum_diff[, tip := rep("Difference between Females and Males", nrow(datum_diff))]
datum_b[, tip := rep("Expected life from birth", nrow(datum_b))]
datum_msch[, tip := rep("Mean years of schooling", nrow(datum_msch))]
datum_expsch[, tip := rep("Expected years of schooling", nrow(datum_expsch))]
datum_eduindex[, tip := rep("Education Index", nrow(datum_eduindex))]

colnamesaremess <- c("iso2c","Country","Result","Continent","tip")
colnames(datum_m) <- colnamesaremess
colnames(datum_f) <- colnamesaremess
colnames(datum_b) <- colnamesaremess
colnames(datum_diff) <- colnamesaremess
colnames(datum_msch) <- colnamesaremess
colnames(datum_expsch) <- colnamesaremess
colnames(datum_eduindex) <- colnamesaremess

setorder(datum_m, -Result)
setorder(datum_f, -Result)
setorder(datum_b, -Result)
setorder(datum_diff, -Result)
datum_diff$Result <- round(datum_diff$Result,2)
setorder(datum_msch, -Result)
setorder(datum_expsch, -Result)
setorder(datum_eduindex, -Result)

datum_life <- rbind(datum_m,datum_f,datum_b)


datum_edu <- rbind(datum_expsch,datum_msch)

datum_diff 

datum_eduindex

# For general statistics part these codes will create the tables.

DT::datatable(datum_b)
DT::datatable(datum_f)
DT::datatable(datum_m)
DT::datatable(datum_diff)
DT::datatable(datum_msch)
DT::datatable(datum_expsch)
DT::datatable(datum_eduindex)





```


